---
title: "Untitled"
output: github_document
---

#### (b) to input the values of a bivariate discrete random variable and the associated probabilities and to obtain the marginal and conditional distributions.Your program should provide a facility to calculate the mean and variance of each distribution, and to plot the pdf and cdf. Do validity checks that the probabilities are in the interval [0, 1], and that they sum to one.
```{r}
# This will ask the user for the title of the distribution
distribution_title <- readline(prompt="Enter the title of the distribution: ")


# This will ask the user if how many sample space or elements there will be in x-axis
input_counter_x <- as.integer(readline(prompt="Enter the number of sample space in X-axis: "))

# Vector or list where the name of each x-value will be stored (EXAMPLE NAMES: Low, Medium, High)
sample_space_x <- c()

# This will ask the user for the name of each x-value
for (i in 1:input_counter_x) {
  ss_x <- readline(prompt="Enter the sample space: ")
  sample_space_x <- c(sample_space_x, ss_x)
}


# This will ask the user if how many sample space or elements there will be in y-axis
input_counter_y <- as.integer(readline(prompt="Enter the number of sample space in Y-axis: "))

# Vector or list where the name of each y-value will be stored (EXAMPLE NAMES: Low, Medium, High)
sample_space_y <- c()

# This will ask the user for the name of each y-value
for (i in 1:input_counter_y) {
  ss_y <- readline(prompt="Enter the sample space: ")
  sample_space_y <- c(sample_space_y, ss_y)
}


# Vector or list where all the variables from x-axis and y-axis will be stored
probabilities_b <- numeric()

# This will ask the user for each joint probabality, depending on the number of x- and y-axis
input_counter <- input_counter_x*input_counter_y
for (i in 1:input_counter) {
  if (i < input_counter) {
    repeat {
      ss_probability <- as.numeric(readline(prompt="Enter its corresponding probability (NOTE: Enter variables BYROW): "))
      
      # Check if the probability is within the valid range
      if (ss_probability >= 0 && ss_probability <= 1) {
        probabilities_b <- c(probabilities_b, ss_probability)
        print(paste0("Initial probability sum: ", sum(probabilities_b)))
        break
      } else {
        cat("The probability must be between 0 and 1. Please enter a valid probability.\n")
      }
    }
  } else {
    remaining_probability <- 1 - sum(probabilities_b)
    cat("Remaining probability added: ", remaining_probability)
    
    # Check if the remaining probability is within the valid range before adding
    if (remaining_probability >= 0 && remaining_probability <= 1) {
      probabilities_b <- c(probabilities_b, remaining_probability)
    } else {
      cat("Error: Sum of entered probabilities exceeds 1. Please check your inputs.\n")
      # Optionally, handle the error (e.g., reset the last entry, adjust the entire vector, etc.)
    }
  }
}
```



##### Matrix of the Bivariate Distribution:
```{r}
distribution <- matrix(probabilities_b, input_counter_x, input_counter_y, byrow = TRUE)
rownames(distribution) <- sample_space_y
colnames(distribution) <- sample_space_x
distribution
```

##### MARGINAL DISTRIBUTIONS
Marginal Distribution along the X-axis (columns):
```{r}
marginal_distribution_x <- colSums(distribution)
marginal_distribution_x
```
Marginal Distribution along the Y-axis (rows):
```{r}
marginal_distribution_y <- rowSums(distribution)
print("Marginal Distribution along the Y-axis (rows):")
print(marginal_distribution_y)
```

##### CONDITIONAL DISTRIBUTIONS
Conditional Distribution of X given Y
P(X=L|Y=L)
```{r}
conditional_distribution_x_given_y <- sweep(distribution, 1, marginal_distribution_y, "/")
print("Conditional Distribution of X given Y:")
print(conditional_distribution_x_given_y)

```
Conditional Distribution of Y given X
```{r}
conditional_distribution_y_given_x <- sweep(distribution, 2, marginal_distribution_x, "/")
print("Conditional Distribution of Y given X:")
print(conditional_distribution_y_given_x)
```

##### PLOTTING THE PROBABILITY MASS FUNCTION
```{r}
library(ggplot2)

prob_df <- as.data.frame(as.table(distribution))

# Plotting
ggplot(prob_df, aes(Var1, Var2, fill = Freq)) + 
  geom_tile() + 
  geom_text(aes(label = sprintf("%.2f", Freq)), color = "black") +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(x = "Variable X", y = "Variable Y", fill = "Probability") +
  theme_minimal() +
  ggtitle(distribution_title)
```

##### PLOTTING THE CUMULATIVE DISTRIBUTION FUNCTION
```{r}
x_axis_labels <- c()
for (i in 1:length(sample_space_y)) {
  for (j in 1:length(sample_space_x)) {
    x_y_label <- paste(sample_space_y[i], sample_space_x[j], sep = "_")
    x_axis_labels <- c(x_axis_labels, x_y_label)
  }
}
x_values <- 1:length(x_axis_labels)

y <- cumsum(probabilities_b)

plot(x_values, y, type = "s",
     xlab = "Sample Space", ylab = "Cumulative probability (P(X<=x))",
     main = distribution_title, xaxt = 'n')

# Add custom x-axis labels
axis(1, at = x_values, labels = x_axis_labels, las = 1)
```

















