---
title: "Untitled"
output: github_document
---

#### (b) to input the values of a bivariate discrete random variable and the associated probabilities and to obtain the marginal and conditional distributions.Your program should provide a facility to calculate the mean and variance of each distribution, and to plot the pdf and cdf. Do validity checks that the probabilities are in the interval [0, 1], and that they sum to one.
```{r}
distribution_title <- readline(prompt="Enter the title of the distribution: ")


input_counter_x <- as.integer(readline(prompt="Enter the number of sample space in X-axis: "))
# Sample Space Vector X
sample_space_x <- c()
for (i in 1:input_counter_x) {
  ss <- readline(prompt="Enter the sample space: ")
  sample_space_x <- c(sample_space_x, ss)
}

input_counter_y <- as.integer(readline(prompt="Enter the number of sample space in Y-axis: "))
sample_space_y <- c()
for (i in 1:input_counter_y) {
  ss <- readline(prompt="Enter the sample space: ")
  sample_space_y <- c(sample_space_y, ss)
}

input_counter <- input_counter_x*input_counter_y

# Probability Vector
probabilities_b <- numeric()



for (i in 1:input_counter) {
  if (i < input_counter) {
    repeat {
      ss_probability <- as.numeric(readline(prompt="Enter its corresponding probability: "))
      
      # Check if the probability is within the valid range
      if (ss_probability >= 0 && ss_probability <= 1) {
        probabilities_b <- c(probabilities_b, ss_probability)
        print(paste0("Initial probability sum: ", sum(probabilities_b)))
        break
      } else {
        cat("The probability must be between 0 and 1. Please enter a valid probability.\n")
      }
    }
  } else {
    remaining_probability <- 1 - sum(probabilities_b)
    cat("Remaining probability added: ", remaining_probability)
    
    # Check if the remaining probability is within the valid range before adding
    if (remaining_probability >= 0 && remaining_probability <= 1) {
      probabilities_b <- c(probabilities_b, remaining_probability)
    } else {
      cat("Error: Sum of entered probabilities exceeds 1. Please check your inputs.\n")
      # Optionally, handle the error (e.g., reset the last entry, adjust the entire vector, etc.)
    }
  }
}
```

##### SAMPLE SPACE:
```{r}
distribution <- matrix(probabilities_b, input_counter_y, input_counter_x)
rownames(distribution) <- sample_space_x
colnames(distribution) <- sample_space_y
distribution
```

##### MARGINAL DISTRIBUTIONS
Low Amount of Time on Phone
```{r}
x_1 <- sum(distribution[1:3, 1])
x_1
```
Medium Amount of Time on Phone
```{r}
x_2 <- sum(distribution[1:3, 2])
x_2
```
High Amount of Time on Phone
```{r}
x_3 <- sum(distribution[1:3, 3])
x_3
```
Low Amount of Time on Reading
```{r}
y_1 <- sum(distribution[1, 1:3])
y_1
```
Medium Amount of Time on Reading
```{r}
y_2 <- sum(distribution[2, 1:3])
y_2
```
High Amount of Time on Reading
```{r}
y_3 <- sum(distribution[3, 1:3])
y_3
```

##### CONDITIONAL DISTRIBUTIONS
##### Y = L
P(X=L|Y=L)
```{r}
round((distribution[1, 1]) / y_1, 2)
```
P(X=M|Y=L)
```{r}
round((distribution[1, 2]) / y_1, 2)
```
P(X=H|Y=L)
```{r}
round((distribution[1, 3]) / y_1, 2)
```
##### Y = M
P(X=L|Y=M)
```{r}
round((distribution[2, 1]) / y_2, 2)
```
P(X=M|Y=M)
```{r}
round((distribution[2, 2]) / y_2, 2)
```
P(X=H|Y=M)
```{r}
round((distribution[2, 3]) / y_2, 2)
```
##### Y = H
P(X=L|Y=H)
```{r}
round((distribution[3, 1]) / y_3, 2)
```
P(X=L|Y=H)
```{r}
round((distribution[3, 2]) / y_3, 2)
```
P(X=L|Y=H)
```{r}
round((distribution[3, 3]) / y_3, 2)
```
##### X = L
P(Y=L|X=L)
```{r}
round((distribution[1, 1]) / x_1, 2)
```
P(Y=M|X=L)
```{r}
round((distribution[2, 1]) / x_1, 2)
```
P(Y=H|X=L)
```{r}
round((distribution[3, 1]) / x_1, 2)
```
##### X = M
P(Y=L|X=M)
```{r}
round((distribution[1, 2]) / x_2, 2)
```
P(Y=M|X=M)
```{r}
round((distribution[2, 2]) / x_2, 2)
```
P(Y=H|X=M)
```{r}
round((distribution[3, 2]) / x_2, 2)
```
##### X = H
P(Y=L|X=H)
```{r}
round((distribution[1, 3]) / x_3, 2)
```
##### X = H
P(Y=M|X=H)
```{r}
round((distribution[2, 3]) / x_3, 2)
```
P(Y=H|X=H)
```{r}
round((distribution[3, 3]) / x_3, 2)
```





