---
title: "Untitled"
output: github_document
---


### 2. With your own computing experience, develop a front end to R that allows the user
#### (a) to input the values of a univariate discrete random variable and the associated probabilities and to obtain the mean and variance. Do validity checks that the probabilities are in the interval [0, 1], and that they sum to one.
```{r setup}
# This will ask the user for the title of the distribution
distribution_title <- readline(prompt="Enter the title of the distribution: ")


# This will ask the user if how many sample space or elements there will be
input_counter <- as.integer(readline(prompt="Enter the number of sample space: "))

# Vector or list where each sample space or the name of each probability will be stored (EXAMPLE: Head or Tails)
sample_space <- c()

# Vector or list where all the probabilities itself will be stored (EXAMPLE: 0.6, 0.4)
probabilities_u <- numeric()

# This will ask the user for each sample space and its corresponding probability
# TAKE NOTE that the last corresponding probability does not need to be entered, it will be automatic
for (i in 1:input_counter) {
  ss <- as.integer(readline(prompt="Enter the sample space: "))
  sample_space <- c(sample_space, ss)
  
  if (i < input_counter) {
    repeat {
      ss_probability <- as.numeric(readline(prompt="Enter its corresponding probability: "))
      
      # Check if the probability is within the valid range
      if (ss_probability >= 0 && ss_probability <= 1) {
        probabilities_u <- c(probabilities_u, ss_probability)
        print(paste0("Initial probability sum: ", sum(probabilities_u)))
        break
      } else {
        cat("The probability must be between 0 and 1. Please enter a valid probability.\n")
      }
    }
  } else {
    remaining_probability <- 1 - sum(probabilities_u)
    
    # Check if the remaining probability is within the valid range before adding
    if (remaining_probability >= 0 && remaining_probability <= 1) {
      probabilities_u <- c(probabilities_u, remaining_probability)
    } else {
      cat("Error: Sum of entered probabilities exceeds 1. Please check your inputs.\n")
      # Optionally, handle the error (e.g., reset the last entry, adjust the entire vector, etc.)
    }
  }
}
```



##### SAMPLE SPACE:
```{r}
sample_space
```

##### CORRESPONDING PROBABILITIES:
```{r}
probabilities_u
```


##### MEAN:
```{r}
distribution_mean <- sum(sample_space*probabilities_u)
distribution_mean
```


##### VARIANCE:
```{r}
distribution_variance <- sum((sample_space - distribution_mean)^2 * probabilities_u)
distribution_variance
```


##### PLOTTING THE PROBABILITY MASS FUNCTION
```{r}
barplot(probabilities_u, names.arg = sample_space, col = "blue",
        xlab = "Sample Space", ylab = "Probability",
        main = distribution_title)
```


##### PLOTTING THE CUMULATIVE DISTRIBUTION FUNCTION
```{r}
x <- 1:length(sample_space)
y_1 <- cumsum(probabilities_u)


plot(x, y_1, type = "s",
     xlab = "Sample Space", ylab = "Cumulative probability (P(X<=x))",
     main = distribution_title)

# Add custom x-axis labels
axis(1, at = x, labels = sample_space, las = 1)
```












